Modelling growth of urban firm networks

### 2022-10-19 communications/CCS2022_UrbanSys/slides.tex
slides_urbansysCCS22

\title
[A generative model for urban firm networks]{A generative model for urban firm networks}
\author[Raimbault \and Zdanowska \and Arcaute] {J.~Raimbault$^{1,2,3}$ \and N. Zdanowska$^{1,3}\ast$ \and E. Arcaute$^{1}$\\\medskip
$^{\ast}$\texttt{n.zdanowska@ucl.ac.uk}
}

\institute[UCL]{$^{1}$Center for Advanced Spatial Analysis, University College London\\
$^{2}$UPS CNRS 3611 Complex Systems Institute Paris\\
$^{3}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}

\date[24/09/2020]{NetSci 2020\\
Session 12F: Diffusion\\
September 24th, 2020
}

@@

@@

General framework:

@@

@@

Assumptions:

@@

\begin{itemize}
	\item Indirect inference on processes
	\item Includes path-dependency
\end{itemize}

@@

Cities characterized by economic size $E_i$ (GDP) and economic structure $S_{ik}$ (probability distribution of firms within $K$ sectors)

@@

Starting from an initial network, at each time step, add a fixed number of links randomly, following a probability as a generalized Cobb-Douglas function

@@

\sframe{Statistical models}{

@@

\resizebox{\linewidth}{!}{
\begin{tabular}{|l|c|c|c|c|c|}
\hline
Model  & (1) & (2) & (3) & (4) & (5 - Poisson GLM) \\ 
\hline
$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
Countries &    &  28.3\% &   &  26.5\% & 100\% \\
\hline
$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
AIC &        44304   &  43578  &  42131  & 41917 &   \\
\hline
\end{tabular}
}

@@

\begin{enumerate} 
    \item Simulation on synthetic systems of cities \cite{raimbault2019space}: generation of a continent-scale urban system with stylized order of magnitude corresponding to Europe (urban hierarchy, countries, industrial sector distributions - \hyperlink{syntheticsetup}{\beamerbutton{Details}})
    \bigskip
    \item Simulation on the European system of cities (AMADEUS data at FUA level)
\end{enumerate}

@@

    % One factor sampling results
    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID

@@

    \textit{(Left) Internationalization index decreases exponentially with gravity decay; (Right) Correlation between city weighted degree and size. Both plots show a transition from a local to a global regime.}

@@

% \sframe{Effect of sector proximity}{

    % One factor sampling results
    %  20190924_162740_ONEFACTOR_REPLICATIONS_SYNTHETIC_GRID

     % \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gammaSectors_errorbars.png}
    % \includegraphics[width=0.48\textwidth]{../../figuresraw/metropolisation-gammaSectors_errorbars.png}

    % \medskip
    % \footnotesize

    % \textit{(Left) Internationalization varies linearly with sector proximity $\gamma_S$; (Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important}

% }

% \sframe{Grid exploration: internationalization}{

    % \begin{center}
    % \includegraphics[width=\textwidth]{../../figures/Fig5.png}
    % \end{center}

    % \medskip

    % \footnotesize

    % \textit{The transition as a function of interaction range % % depends on the influence of origin size $\gamma_F$; sector % proximity $\gamma_S$ plays a role only for a large influence of the origin.}

% }

@@

\sframe{Calibration on the European urban network}{

    \begin{center}
    	\includegraphics[height=0.75\textheight]{../../figures/Fig8.png}

@@

@@

\textit{Although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be
expected to have a detrimental effect on UK economy due to these foreign ownership links.}

@@

\textbf{Practical application}

@@

$\rightarrow$ co-evolution model (evolution of city sizes) \cite{raimbault2018modeling};

@@

\begin{enumerate}
    \item Generate $N \textrm{ = } 700$ cities with size following a power law $E_i \textrm{ = } E_0 \cdot i^{-\alpha}$ with $E_0 \textrm{ = } 10^{11}$ and $\alpha \textrm{ = } 1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
    \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
    \item Create countries with k-means clustering ($C \textrm{ = } 30$)
    \item Distribute sectors such that (i) smaller cities are more specialized and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest
\end{enumerate}

@@

\textit{Impact of urban hierarchy on model indicators in synthetic systems of cities.}

>>>>

\title{Modelling growth of urban firm networks}
\author{J.~Raimbault$^{1,2,3,4,\ast}$ \and N. Zdanowska$^{5,2,4}$ \and E. Arcaute$^{2}$\\
$\ast$\texttt{juste.raimbault@ign.fr}
}

\institute{$^{1}$LASTIG, Univ Gustave Eiffel, IGN-ENSG\\
$^{2}$CASA, UCL\\
$^{3}$UPS CNRS 3611 ISC-PIF\\
$^{4}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
$^{5}$Luxembourg Institute of Socio-Economic Research
}

\date{CCS 2022\\
UrbanSys Satellite\\
19/10/2022
}

@@

% extended abstract
% The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks – in particular urban firm networks –, is essential for the economic resilience of urban systems. We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the source and target nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distances between urban areas. These are combined into a generalised spatial interaction model giving the probability to add links between urban areas, used in an iterative manner to generate the network from an initial configuration.
%An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation. We construct a directed network between functional urban areas in Europe, based on the AMADEUS data which is a well covering dataset describing ownership links between companies and their subsidiaries. The links between firms are aggregated at the urban area level and economic size of areas are computed by aggregating turnovers of firms. A community detection analysis provides a community structure which is partly geographical but also captures some known economic collaboration areas (see Fig. 1). Statistical models (various specifications of spatial interaction models) assess the role of various factors to be included in the simulation model.
%We then simulate network growth on synthetic systems of cities with the generative model. Examples of generated networks are shown in Fig. 1. Global sensitivity analysis provides the influence of parameters on model outputs characterising network structure. A grid sampling of the parameter space unveils stylised facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range.
%We calibrate the model on the European network, with a real geographical setup and economic sector composition obtained from the empirical network data. Calibration is done on two objectives, in order to reproduce the real network in terms of link weights (mean squared error on weights and mean squared error on logarithms of weights), with a genetic algorithm for bi-objective optimisation (NSGA2). Considering the common fitting objective, the simulation model outperforms statistical models. This furthermore shows a strong role of path-dependency for model performance.
%The model is finally applied to stylised scenarios, where economic shocks modify spatial interaction parameters after a fixed number of time steps. This opens up to the formulation of public policies recommendations to deal with exogenous shocks such as economic crises or potential lock-downs of countries.
% Fig 2, 3
% Figure 1: (Left) Empirical analysis of the AMADEUS data: urban area size corresponds to economic size, while color gives the community obtained with a Louvain modularity optimisation algorithm; (Right) Example of simulated networks: top row in a synthetic system with long-range interactions (left) and short range interactions (right), bottom row with same parameter values but in the real setting.

@@

%General framework:
\textit{Role of firm networks in urban systems:}

@@

\textit{Relevance of dynamics urban network evolution models:}

@@

Main geographical processes considered:

@@

Advantages of the model introduced:

\begin{itemize}
	\item Indirect inference on processes
	\item Multiple complementary factors driving inter-urban links
	\item Includes path-dependency
\end{itemize}

@@

Cities characterised by economic size $E_i$ (GDP) and economic structure $S_{ik}$ (probability distribution of firms within $K$ sectors)

@@

Starting from an initial network, at each time step, add a fixed number of links randomly, following a probability as a generalised Cobb-Douglas function

@@

\begin{frame}[label=statmain]

\frametitle{Statistical models}

@@

\resizebox{\linewidth}{!}{
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 & \multicolumn{2}{|c|}{$d_0$} & \multicolumn{2}{|c|}{$c_0$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
 & F & T & F & T & F & T & F & T & F & T & F & T \\
 \hline
Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.008 & $5\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
Metropolisation & 0.02 & 0.04 & 0.02 & 0.04 & 0.02 & 0.3 & 0.002 & 0.01 & 0.2 & 0.7 & 0.2 & 0.7 \\
Modularity & 0.3 & 0.4 & 0.6 & 0.6 & $5\cdot 10^{-4}$ & 0.02 & 0.002 & 0.02 & 0.007 & 0.03 & 0.003 & 0.03 \\
Avg. com. size & 0.004 & 0.1 & 0.01 & 0.1 & 0.003 & 0.07 & 0.001 & 0.04 & 0.3 & 0.6 & 0.3 & 0.6 \\
Degree entropy & 0.003 & 0.03 & 0.002 & 0.03 & 0.008 & 0.04 & 0.003 & 0.03 & 0.5 & 0.6 & 0.5 & 0.6 \\
Weight entropy & 0.04 & 0.2 & 0.03 & 0.2 & 0.002 & 0.1 & 0.001 & 0.1 & 0.5 & 0.6 & 0.5 & 0.6 \\\hline
\end{tabular}
}

Zero-inflated and Hurdle models: \hyperlink{zeroinfl}{\beamerbutton{Details}}

@@

\begin{enumerate} 
    \item Simulation on synthetic systems of cities \cite{raimbault2019space}: generation of a continent-scale urban system with stylized order of magnitude corresponding to Europe (urban hierarchy, countries, industrial sector distributions - \hyperlink{syntheticsetup}{\beamerbutton{Details}})
    \bigskip
    \item Simulation on the European system of cities, setup with AMADEUS data at FUA level (\hyperlink{realsetup}{\beamerbutton{Details}})
\end{enumerate}

@@

@@

    \textit{(Left) Internationalisation index decreases exponentially with gravity decay; (Right) metropolisation also shows a progressive transition from a local to a global regime.}

@@

\sframe{Effect of sector proximity}{

     \includegraphics[width=0.48\textwidth]{../../figuresraw/internationalisation-gammaSectors_errorbars.png}
    \includegraphics[width=0.48\textwidth]{../../figuresraw/metropolisation-gammaSectors_errorbars.png}

    \medskip
    \footnotesize

    \textit{(Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Right) metropolisation witnesses an intermediate regime where size is the most important}

\sframe{Grid exploration: internationalisation}{

     \begin{center}
     \includegraphics[width=\textwidth]{../../figures/Fig5.png}
     \end{center}

     \textit{The transition as a function of interaction range depends on the influence of origin size $\gamma_F$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.}

@@

\textit{Impact of urban hierarchy on model indicators in synthetic systems of cities: existence of a maximum for metropolisation at intermediate hierarchies.}

\begin{frame}[label=calibmain]

\frametitle{Calibration on the European urban network}

    \begin{center}
    	\includegraphics[height=0.7\textheight]{../../figures/Fig8.png}

@@

    % weird point out of pareto front: ?

   \hyperlink{calibration}{\beamerbutton{Details}} on the calibration procedure

@@

\textit{Although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be
expected to have a detrimental effect on UK economy due to these foreign ownership links.}

@@

\textbf{Practical applications}

@@

$\rightarrow$ co-evolution model (evolution of city sizes) \cite{raimbault2021modeling};

@@

\begin{enumerate}
    \item Generate $N \textrm{ = } 700$ cities with size following a power law $E_i \textrm{ = } E_0 \cdot i^{-\alpha}$ with $E_0 \textrm{ = } 10^{11}$ and $\alpha \textrm{ = } 1.1$ (computed on Europe for GDP with cities larger than 50.000 inhabitants)
    \item Distribute them randomly in space (\cite{simini2019testing} vs \cite{banos2011christaller})
    \item Create countries with k-means clustering ($C \textrm{ = } 30$)
    \item Distribute sectors such that (i) smaller cities are more specialised and (ii) larger cities are more knowledge-based, with a one dimensional axis to position sectors $1/K \ldots 1$ where the density $f\left(k\right)$ follows a log-normal with $\left(\mu,\sigma\right)$ such that $\sqrt{\Var f} \textrm{ = } K/2$ for the largest, $\sqrt{\Var f} \textrm{ = } 1/K$ for the smallest % equations? too much -> refer to paper
\end{enumerate}

\begin{frame}[label=realsetup]

\frametitle{Real configuration setup}

% sociocultural distance

Urban areas (FUAs) positions, economic size initialised from the GHSL database, sector composition from the Amadeus database using classification of firms and aggregating

Initial dummy network ($w_0 = 1$)

\textbf{Socio-cultural distance} $c_{ij}$: fixed effects coefficients

% Q: how are NAs handled?: infinite: ~ - anyway not used in calibration

\begin{tabular}{ccccccc}
   Min. & 1st Qu. &  Median &  Mean &  3rd Qu. &  Max. &  NA's \\
 -2.327 &  1.440 &  2.451 &  2.538 &  3.447 & 11.797 &  339 
\end{tabular}

@@

\begin{frame}[label=zeroinfl]

\frametitle{Zero-inflated and Hurdle statistical models}

Accounting for zeros with specific models: same qualitative results

\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{|l|c|c|c|c|}
\hline
Model  & \multicolumn{2}{|c|}{Zero-inflated} & \multicolumn{2}{|c|}{Hurdle} \\ 
 & Count & Zero-infl. & Count & Zero-hurdle \\
\hline
$\log(d_{ij})$ &    -0.28*** (7e-4) &   0.88*** (0.01)      &   -0.29*** (8e-6)   &  -0.75*** (0.01)    \\
$\log(T_i)$ & 0.74*** (4e-6)  & -0.74*** (7e-3)       &   0.74***  (5e-6)     &  0.67*** (5e-3)   \\
$\log(T_j)$ & 0.64*** (4e-6)  & -0.52*** (7e-3)      &    0.64***  (4e-6)     &  0.48*** (5e-3)    \\
$\log(s_{ij})$ & 0.46*** (2e-5)   &  -0.33*** (3e-2)  &  0.45*** (2e-5)       &  0.27*** (0.02)    \\
\hline
$R^2$ &   \multicolumn{2}{|c|}{0.1607}     &  \multicolumn{2}{|c|}{0.1661}  \\
%MSE & & \\
AIC &     \multicolumn{2}{|c|}{5.85e10}    &   \multicolumn{2}{|c|}{5.85e10}  \\
\hline
\end{tabular}
}
\end{center}

\hyperlink{statmain}{\beamerbutton{Back}}

\begin{frame}[label=calibration]

\frametitle{Model calibration}

\textbf{To avoid overfitting:} $c_0 = 0$ to deactivate the mechanism

\textbf{Calibration objectives:}

\[
\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2
\]

\[
\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)
\]

\textbf{On macro indicators?} Too much solutions, equifinality

\textbf{Rescale flows:} $w_0$ is arbitrary, so is optimised with rescaling:

\[
\varepsilon_L (k_0) = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \left(k_0 \cdot \hat{w}_{ij} \right) \right)^2
\]

what yields

\[
k_0 = \exp \left[\frac{1}{N^2} \sum_{i,j} \log w_{ij} - \frac{1}{N^2} \sum_{i,j} \log \hat{w}_{ij} \right]
\]

\hyperlink{calibmain}{\beamerbutton{Back}}
